### 2015-10-17 paper.tex
section 4 and 5

%\documentclass{sig-alternate}
\documentclass{acm_proc_article-sp}

\usepackage[utf8]{inputenc}
\usepackage[hyphens]{url}
\usepackage[pdftex,urlcolor=black,colorlinks=true,linkcolor=black,citecolor=black]{hyperref}
\def\sectionautorefname{Section}
\def\subsectionautorefname{Subsection}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{gensymb}
% todo macro
\usepackage{color}
\newcommand{\todo}[1]{\noindent\textcolor{red}{{\bf \{TODO}: #1{\bf \}}}}

\title{From Freebase to Wikidata: The Great Migration}

\numberofauthors{5}
\author{
\alignauthor
Thomas Pellissier Tanon\titlenote{The author was an intern at Google when the majority of the work was done.}
   \affaddr{Google, San Francisco, USA}\\   
   \email{\mbox{thomas@pellissier-tanon.fr}}
\alignauthor
Denny Vrandečić\\
   \affaddr{Google, San Francisco, USA}\\   
   \email{vrandecic@google.com}
\alignauthor
Sebastian Schaffert\\
   \affaddr{Google, Zurich, Switzerland}\\   
   \email{schaffert@google.com}
\and
\alignauthor
Thomas Steiner\\
   \affaddr{Google, Hamburg, Germany}\\   
   \email{tomac@google.com}
\alignauthor
Lydia Pintscher\\
   \affaddr{Wikimedia, Berlin, Germany}\\   
   \email{lydia@pintscher.de}
}
\begin{document}

\maketitle

\begin{abstract}
Collaborative knowledge bases that make their data freely available in a~machine-readable form
are central for the data strategy of many projects and organizations.
The two major collaborative knowledge bases are Google's Freebase and Wikimedia's Wikidata.
Due to the success of Wikidata, Google decided in 2014 to offer the content of Freebase
to the Wikidata community.
In this paper, we describe the ongoing transfer efforts and data mapping challenges,
and provide an analysis of the effort so far.
We describe the \emph{Primary Sources Tool}, which aims to facilitate
this and future data migrations.
Throughout the migration, we have gained deep insights into both Wikidata and Freebase,
and share and discuss detailed statistics on both knowledge bases. 
\end{abstract}

% \category{CR number}{category}{sub-category}[optional subject descriptor]
\category{H.3.5}{Information Storage and Retrieval}{Online Information Systems}
\category{H.5.3}{Information Interfaces}{Group and Organization Interfaces}[Web-based interactions]
\category{I.2.4}{Artificial Intelligence}{Knowledge Representation}
\category{K.4.3}{Computers and Society}{Organizational Impacts}[Com\-puter-supported collaborative work]

\terms{Human Factors, Documentation}

\keywords{Crowdsourcing systems, Semantic Web, Wikidata, Freebase}

\section{Introduction}

Large Web-based knowledge bases that make their data available
under free licenses in a~machine-readable form
have become central for the data strategy of many projects and organizations.
They find applications in areas as diverse as Web search,
natural language annotation, or translation.
Exemplary works are those of West \emph{et~al.}, who have proposed~\cite{west2014kb} a~way to
leverage existing Web search based question-answering technology to
fill in the gaps in knowledge bases in a~targeted way
or those of Gesmundo and Hall~\cite{gesmundo2014kb}, who have trained
a~syntactic parser paradigm that learns from large scale knowledge bases.

One such collaborative knowledge base is Freebase,
publicly launched by Metaweb in 2007 and acquired by Google in 2010.
Another example is Wikidata, a~collaborative knowledge base developed by
Wikimedia Deutschland since 2012 and operated by the Wikimedia Foundation.
Due to the success of Wikidata, Google announced in 2014 to shut down Freebase and
help the community with the transfer of Freebase's content to Wikidata~\cite{freebase2014shutdown}.

Moving data between two knowledge bases that do not share a~similar design
is usually a~problematic task and requires the careful mapping between their structures.
The migration from Freebase to Wikidata was no exception to this rule:
we encountered a~number of to-be-expected \emph{structural} challenges.
However, even more demanding was the \emph{cultural} difference between the two involved communities.
The Freebase and Wikidata communities have a~very different background,
subtly different goals and understandings of their tasks,
and different requirements regarding their data.

In this paper, we describe how we support the Wikidata community with
the migration of Freebase content to Wikidata.
We programmed the \emph{Primary Sources Tool} and
released it as open source software to facilitate the transfer of the Freebase dataset.
The \emph{Primary Sources Tool} is also developed with an eye on
future data migrations from other sources.
We created and published mappings of the Freebase dataset.
Throughout the migration, we have gained deep insights into both Wikidata and Freebase,
and share and discuss detailed statistics on both knowledge bases.

The paper is structured as follows.
First, we introduce Freebase and Wikidata in \autoref{sec:background}.
Then we describe our methodology and the metrics used to measure the migration
in \autoref{sec:challenges-of-the-migration}.
In order to support the migration, we have developed a~set of open source tools
that we present in \autoref{sec:primary-sources-tool}.
We then show the results of the migration
and discuss these statistics in \autoref{sec:statistics-of-the-migration}.
Before we close, we report on related work in \autoref{sec:related-work},
followed by an outlook at proposed next steps
and a~conclusion in \autoref{sec:future-work-and-conclusion}.

\section{Background}\label{sec:background}

\subsection{Freebase}

Freebase\footnote{Freebase: \url{https://www.freebase.com}} is an open and
collaborative knowledge base publicly launched in 2007 by Metaweb and 
acquired in 2010 by Google~\cite{freebase}.
It was used as the open core of the Google Knowledge Graph,
and has found many use cases outside of Google.
Due to the success of Wikidata,
Google announced in 2014 to close Freebase and offer most of its content to the
Wikidata community~\cite{freebase2014shutdown,freebase-end}.

Freebase is built on the notions of \emph{objects}, \emph{facts}, \emph{types}, and \emph{properties}.
Each Freebase object has a~stable identifier called a~\emph{``mid''} (for Machine ID),
one or more types, and uses properties from these types in order to provide facts.
For example, the Freebase object for Barack Obama has the mid \texttt{/m/02mjmr}
and the type \texttt{/government/us\_president} (among others) that allows the entity to have
a fact with the property \texttt{/government/us\_president/presidency\_number} \linebreak
and the literal integer ``44'' as the value.

Freebase uses Compound Value Types (CVTs) to represent
n-ary relations with $n>2$, \emph{e.g.}, values like geographic coordinates,
political positions held with a~start and an end date (see \autoref{fig:cvt-obama} for an example),
or actors playing a~character in a~movie.
CVT values are just objects, \emph{i.e.}, they have a~mid and can have types
(although they usually have only the compound value type itself).
Most non-CVT objects are called \emph{topics} in order to discern them from CVTs.

\begin{figure}[!htbp]
  \centering
  \includegraphics[trim=0cm 3cm 0cm 5cm, width=8.45 cm]{img/freebase-cvt-obama.pdf}
  \caption{Exemplary Compound Value Type (CVT) in Freebase: Presidency of Barack Obama.}
  % Source: https://docs.google.com/document/d/1ELrBtxxs4ghhe7FloTUnhvpD3f7RFczT7UfxahAfjgE/edit
  \label{fig:cvt-obama}
\end{figure}

The content of Freebase has been partially imported from various sources
such as Wikipedia~\cite{wikipedia-works} or
the license-compa\-tible part of MusicBrainz~\cite{musicbrainz}.
Over the years, the Freebase community and Google have maintained the knowledge base.
When Freebase was turned read-only on March~31, 2015,
it counted more than 3~billion facts about almost 50M~entities.
The data from Freebase is published as an N-Triples dump in RDF~\cite{rdf} under
the Creative Commons CC-BY license.\footnote{Dump: \url{https://developers.google.com/freebase/data}}

\subsection{Wikidata}

Wikidata\footnote{Wikidata: \url{https://www.wikidata.org}}
is a~collaborative knowledge base
launched in October 2012 and hosted by the Wikimedia Foundation~\cite{vrandevcic2014wikidata}.
Its community has been growing quickly, and as of mid 2015,
the community comprises about 6,000 active contributors~\cite{wikidatastats}.

Wikidata's data model relies on the notions of \emph{item} and \emph{statement}.
An item represents an~entity, has a~stable identifier called ``qid'', and may have labels,
descriptions, and aliases in multiple languages; further statements and links to pages
about the entity in other Wikimedia projects---most prominently Wikipedia.
Contrary to Freebase, Wikidata statements do not aim to encode true facts,
but \emph{claims} from different sources, which can also contradict each other.

A~statement is composed of a~claim and zero or more references for this claim.
The claim itself is composed of one main property--value pair that encodes
the main (claimed) fact like ``population is 8,173,900'' and optional qualifiers
to add information about it like ``as of June 2012''.
\autoref{fig:statement} illustrates the used terminology with an example.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=8.45 cm]{img/Wikidata_statement.pdf}
  \caption{Wikidata terminology illustrated with an example (Source:
	\url{https://commons.wikimedia.org/wiki/File:Wikidata_statement.svg}).}
  \label{fig:statement}
\end{figure}

The content of Wikidata is in the public domain under Creative Commons CC0.
As of September 2015, Wikidata counted about 70M~statements on 18M~entities.
A~more in-depth comparison between Freebase, Wikidata, and other free and open knowledge bases
(DBpedia, YAGO, and OpenCyc) is available in~\cite{farbercomparative} by Färber \emph{et~al.}

\section{Challenges of the migration}\label{sec:challenges-of-the-migration}

During our ongoing efforts with the migration of Freebase to Wikidata,
we were faced with a~number of technical and non-technical challenges
that we will outline in the following.

\subsection{Licensing}
\label{sec:licensing}

The first challenge concerns the licenses under which the datasets are published.
Wikidata is published under the Creative Commons~0 (CC0~1.0)~\cite{cc0} license,
which effectively puts the data into the public domain.
Freebase is published under a~Creative Commons Attribution (CC~BY~2.5) license.
Google does not own the copyright of some parts of the content of the knowledge base,
such as images or long entity descriptions extracted from Wikipedia.

We filtered the Freebase dump by removing this kind of content
before creating a~data dump that Google could relicense under CC0.
This step reduced the number of Freebase facts that could be republished by about 42 million facts
from 3~billion facts in the original corpus, i.e. by about $1.4\%$.

\subsection{References}

The second challenge is that the Wikidata community is very eager to have references
for their statements, \emph{i.e.}, sources that Freebase usually did not store,
except for some specific data like populations and unemployment rates.
For these two specific cases, we have just kept the references from Freebase.
\todo{Thomas, can you expand?}

In order to provide the Wikidata community with references for the facts in Freebase,
we have reused data from the Google
Knowledge Vault~\cite{dong2014knowledge}. Knowledge Vault aims at extracting facts from the Web.

The result of this step was surprising: 
the fact extraction was usually correct, in the sense that the fact was indeed being stated on the page.
Also the fact itself was usually accurate, in the sense that it corresponded to reality.
But the pages the facts were extracted from in many cases did not meet
the Wikidata requirements for reliable references:
they include pages in social networks, reviews on shopping sites, file sharing hosters, \emph{etc.}
It became necessary to filter the references before potential inclusion in Wikidata
by introducing a~domain blacklist.%
\footnote{Domain blacklist:
\url{https://www.wikidata.org/wiki/Wikidata:Primary_sources_tool/URL_blacklist}}

\subsection{Data Quality}
\label{sec:dataquality}

The data quality of Freebase was examined thoroughly by the Wikidata community
and anecdotal evidence indeed confirmed that there were certain challenges.
For example, a~small city in France was assigned an ISO country code,%
\footnote{Freebase topic page of the city of Saint-Martin: \url{http://www.freebase.com/m/03cc0d0}}
then Boston (the city in Massachusetts) had the type \texttt{/people/person},
and several other occurrences of data quality issues were revealed.
In consequence, a~fully automatic upload of all the content from Freebase into Wikidata
did not seem advisable as the expectations of the Wikidata community regarding the
quality of automatically uploaded data are high (and rightly so).
Despite the fact that an automated upload of all the data
would have led to accelerated growth in the short term,
such a~fast increase of available data would certainly have endangered
the sense of ownership of the Wikidata community for its data.
As an alternative approach, we decided to rely on crowdsourced human curation
and created the \emph{Primary Sources Tool},
a~widget that displays Freebase statements
for curation by the contributor that can be added to the currently shown Wikidata item.
The tool will be described in more detail in \autoref{sec:primary-sources-tool}.

\subsection{Long-term maintenance}
\label{sec:longtermmaintenance}

Wikidata's editing community is in charge of the project's content. This includes collecting the
data as well as long-term maintenance of it. Data needs to be updated as the world changes, but also
kept free of vandalism. An increase in the amount of data comes with an increase in maintenance
cost for the community. Simply ingesting the Freebase data would have meant overwhelming the
existing editors and thereby harming the project in the long run. To counter this, a~significant
increase in the amount of data needs to go along with either an increase in the number of editors
or provision of tools to assist the existing editors to allow them to be more effective. With the
Primary Sources Tool, we aim to help grow the community in step with the amount of data. It is
useful as a~means to get more people to edit because it lowers the barrier to entry. It does this
by making it easier to find potential edits. At the same time, it helps make some of the existing
editors more effective by allowing them to add statements with a~single click.

\subsection{Data Topic Mappings}\label{subsec:topic-mappings}

\subsubsection{Merger of Existing Mappings}

The next challenge concerns the data mappings between Freebase topics and properties
and Wikidata items and properties.
Two mappings between Freebase topics and Wikidata items were initially available,
and in addition to those, we worked on further mappings.
A~first one has been created by Google in October 2013%
\footnote{Freebase mappings: \url{http://developers.google.com/freebase/data}}
and is based on Wikipedia links already present in Freebase: if a~Freebase topic and
a~Wikidata item share at least two Wikipedia links, they are assumed to describe the same subject.
The quality of this mapping is considered very good by the Wikidata community,
who thus have decided to use it and further curate and maintain it by hand.
It maps 1.15~million Freebase topics to their corresponding Wikidata items.
A~second actively maintained mapping has been created by Samsung.%
\footnote{Samsung mappings: \url{http://github.com/Samsung/KnowledgeSharingPlatform}}
It is based on the same idea, but matches a~Freebase topic with a~Wikidata item
even if there is only a~single shared Wikipedia link.
The precision of the mapping is naturally lower compared to Google's mapping
as there are more wrong matches (often because there is no clear separation between topics
and disambiguation pages in Wikipedia), however, the recall increases to 4.4~million~links.
As we have shown in \autoref{sec:dataquality},
human curation is required before import anyway,
which is why we have decided to merge the two mapping dumps.
For the 6,000 conflicts that appeared, we decided to prioritize Google's mapping.

\subsubsection{Reconciliation Based on External Database IDs}

To improve the result, we have also done some reconciliation based on third-party database IDs
shared by Freebase and Wikidata, like MusicBrainz~\cite{musicbrainz},
VIAF~\cite{viaf}, \emph{etc.}
With this technique, we were able to match 800,000~external database IDs,
creating an additional mapping for 600,000~topics---%
most of them already in the merged version of Google's and Samsung's mappings.
Eventually, this reconciliation effort resulted in an additional 84,000 mapped items.

We also used Google's Knowledge Graph~\cite{singhal2012} to add 100,000 further mappings
by also matching a~topic and an item if they share a~Wikipedia link with the Knowledge Graph item.
There is some potential to add more data by looking at functional relations
like \textit{``father''} and \textit{``mother''},
but these relations only apply to relatively small sets of data
(like well known families), and the functional assumptions may run into some edge cases.
For example, the \texttt{/people/person/parents} property in Freebase also covers stepparents,
which may create cases where a~person has more than one male or female \texttt{/people/person/parents}.

Eventually, with these four sources (Google mapping, Samsung mapping, external IDs, Knowledge Graph),
we mapped 4.56~million items in total.

\subsubsection{Topic Mapping Completeness Analysis}

Thanks to these efforts, we have already mapped most of the topics that could be mapped automatically,
which is also backed by the comparably small increases gained from the ID reconciliation approach 
and the Knowledge Graph data described in the previous paragraph.
This is no surprise, since the core of both projects was kickstarted by using
Wikipedia as a~repository of topics.

The \emph{mapped} topics have an average number of 13.9~facts,
whereas the topics that were \emph{not} mapped have an average number of only 5.7~facts.
We have mapped the majority of all topics with more than 47~facts
(191K~non mapped topics \textit{vs.} 195K~mapped topics).
This indicates that we have mapped the most important items of Freebase.
\autoref{fig:mapped-not-mapped} provides an overview of mapped and non-mapped items
put in relation to the number of facts per item: note that the Figure is log-scaled
on both axes, so the seemingly small area between the two lines covers a majority of
the topics. It can also be seen that there are a very small number of topics with
a high number of facts which have not been covered (in the lower part of the Figure).
These have all been manually inspected and found to be specific to the way knowledge
in Freebase is built. The vast majority of items with more than ten facts - in the
center of the Figure - has been mapped. The items that have not been mapped -
visible in the upper left corner of the image - are very thin,
i.e. mostly have a very small number of facts and
would not have added much to the knowledge base.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=8.45 cm]{img/facts-topics-mapping.pdf}
  \caption{Number of mapped and not mapped topics with more than $x$ facts (log scale on both axes).}
  \label{fig:mapped-not-mapped}
  %Data sheet: https://docs.google.com/spreadsheets/d/1jYGVIqAmyouHKjvhM6FbpgTD8bb5tkJy9PEMELkuMwo/edit  
\end{figure}

\subsection{Data Property Mapping}

For the mapping between Freebase and Wikidata properties, we have chosen to apply a~manual
approach with the help of the Wikidata community.
Note that a mapping of types is not needed, since Wikidata has no real notion of types
(or, to be more specific, the property \textit{``instance of''} (P31) is just another
property and connects to items without any special status).
With the help of the community, we have been able to quickly map around 360~properties
(note that Wikidata had a~total of about 1,800 properties at the time when this work
has been conducted in the Summer of 2015, and Freebase has around 37,700 properties).
For the mapping of properties where the domain are other topics or literals,
this mapping provides the Wikidata property to use.

For \texttt{/people/person/parents}, we have implemented a~special case
in order to be able to map them to the Wikidata properties
\textit{``father''} (P22) and \textit{``mother''} (P25)
based on the person's gender.

The mapping of CVTs (see \autoref{fig:cvt-obama}) is more complicated,
because it is not possible to have a~\mbox{1-to-1} relationship
between Freebase and Wikidata properties: the CVT is linked to the subject topic by one property
and has properties pointing to its component values, whereas the Wikidata statement
has a~main property value group that is qualified by other such groups.
In order to map a~CVT to a~statement, we have to know which of the CVT's properties
should be used as the main value of the statement, with the others being mapped to qualifiers.
In order to keep a~simple mapping that associates one Wikidata property to one Freebase property,
we map both the property that links the topic to the CVT
and the main property of the CVT to the same Wikidata property. In a second step, we
then use merge these two mappings int a single Wikidata claim, on which we add the
other properties from the CVT.

For example, if we map the CVT shown in \autoref{fig:cvt-obama},
we first use the Wikidata property \textit{``position held''} (P39) to map
\texttt{/politician/government\_positions\_held} and \linebreak
\texttt{/government\_position\_held/office\_position\_or\_title}. \linebreak
Second, for the qualifiers, we use \textit{``start time''} (P580)
for \texttt{/government\_position\_held/from} and \textit{``end time''} (P582)
for \texttt{/government\_position\_held/to}.
With the resulting \linebreak property mapping, the created Wikidata statement
looks like the one presented in \autoref{fig:statement-obama}.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=1\columnwidth]{img/wikidata-statement-obama.pdf}
  \caption{Wikidata statement for a~Freebase Compound Value Type (CVT).
    See \autoref{fig:cvt-obama} for comparison.}
  \label{fig:statement-obama}
\end{figure}

For CVTs that include sources, we map the property that links the CVT
to the source as a~Wikidata reference instead of a~qualifier.
The hardest part then is to actually create the full topic mapping
by considering CVTs and the complex datatypes of Wikidata.
For CVTs, when we have a~triple whose value is a~CVT, we just retrieve it,
map all its component triples (while ignoring ones that cannot be mapped)
and apply the above mechanism to get its main value.
For values, we use the datatype information stored with Wikidata properties
in order to cast values to the right type.
For quantities, we assume that the amounts are precise,
as there is no precision information in Freebase.
For globe coordinates, we have hardcoded the mapping of the \texttt{/location/geocode} CVT
into the globe-coordinate value type of Wikidata
and use the same precision guessing algorithm as the one used by the Wikidata user interface.
For dates, we just map the date and time values
(encoded using XSD types~\cite{campbell2003xmlschema})
to the Wikidata ``time'' datatype. However, we remove dates before 1920 that have
a~precision higher than year because we do not know
if they are relative to the Julian or the Gregorian calendar.
In such cases, we are not able to fill the ``calendar'' value of the Wikidata ``time'' datatype.

\section{Primary Sources Tool}\label{sec:primary-sources-tool}

As outlined earlier, the \emph{Primary Sources Tool} is a~crowdsourced human curation
software solution that displays Freebase statements for verification to the Wikidata contributor
so that these statements can be added to the currently shown Wikidata item.
With just one click, the contributor can reject or approve a~statement,
and, in case of approval, add the statement to Wikidata.
The code of the Primary Sources Tool is openly available
under the terms of the Apache~2.0 license (\url{https://github.com/google/primarysources}).
The tool is deployed as a~so-called gadget,
\emph{i.e.}, as an easily to active add-on feature in Wikidata.
It is independent from Freebase and can easily be reused to import other datasets into Wikidata.

We have implemented an upload API in the back-end
and a~dataset filter in the front-end that allows users
to display only statements from a~selected dataset.
This flexibility is indeed much-wanted, and during the development period of the project
another researcher approached us in order to upload
a~dataset extracted from natural language text.
We expect more datasets to be added to the Primary Sources Tool in the future.

At the time of writing (October, 2015), the tool has been used by more than a~hundred users
who performed about 30,000 approval or rejection actions.
More than 12~million statements have been uploaded in total.
\autoref{fig:primary-sources-tool} shows two screenshots
of the tool that integrates nicely with the Wikidata user interface.
To visualize the migration progress, we have created a~realtime dashboard.%
\footnote{Primary Sources Dashboard:
\url{https://tools.wmflabs.org/wikidata-primary-sources/status.html}}

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{1.0\columnwidth}
        \includegraphics[width=\textwidth]{img/primary-sources.pdf}
        \caption{Incoming references.}
        \label{fig:barack-obama}
    \end{subfigure}
    \begin{subfigure}[b]{1.0\columnwidth}
        \includegraphics[width=\textwidth]{img/wikidata-cause-of-death.pdf}
        \caption{Incoming claims and references.}
        \label{fig:cause-of-death}
    \end{subfigure}
    \caption{Primary Sources Tool user interface. Statements coming from
       Freebase that are not yet available in Wikidata are shown in light blue.
       Users can approve or reject these statements using the ``approve'' or ``reject'' buttons.}
    \label{fig:primary-sources-tool}
\end{figure}

\autoref{fig:architecture} shows the overall architecture of the Primary Sources tool.
In the following, we describe the back-end, the front-end, and the chosen data rollout plan.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.65\columnwidth]{img/architecture.pdf}
    \caption{Primary Sources Tool architecture. The front-end (JavaScript) sends
      HTTP requests to the back-end (C++) to retrieve and update entities or
      statements, stored in a~relational database. Approved statements are in
      addition written by the front-end to the Wikidata REST API.}    
    \label{fig:architecture}
    % Source: https://docs.google.com/a/google.com/drawings/d/1vKAftQVjrso907UtbMXnEOabXwQHDVyyz2ewxcqz3b0/edit?usp=sharing
\end{figure}

\subsection{Back-end}

The objective of the back-end was to have a~REST API allowing us
to serve data efficiently to the front-end of the tool and
providing statistical data about the approvals and rejections done over time.
As the resources of the Wikimedia Foundation are limited,
the back-end was implemented in highly optimized C++
and runs as a~FastCGI application in the lighttpd Web server.

Entities and statements are stored in a~relational database
(first SQLite, later migrated to MariaDB)
and follow the Wikidata data structure with an additional status field
to keep track of statements that either already have been approved or rejected,
or not yet visited.
Statements can be grouped in ``datasets'' (set of statements from the same origin)
and ``uploads'' (set of statements uploaded at the same time) in order to
distinguish data from different sources.\footnote{At the time of this writing,
the back-end contained five different datasets.}

The REST API of the back-end supports the following main types of requests:

\begin{itemize}
  \setlength\itemsep{0em}
  \item Get an entity (\emph{i.e.}, its statements) by Wikidata QID
  (\emph{e.g.}, \verb|GET /entities/Q2529789|).
  \item Get all entities with unapproved (or approved or rejected) statements
  (\verb|/entities/all|).
  \item Get a~random entity with unapproved (or approved or rejected) statements
  (\verb|/entities/any|).
  \item Get a~statement by database ID (\verb|GET /statements/1|), get all
    statements (\verb|GET /statements/all|), or get a~list of
  random statements (\verb|GET /statements/any|).
  \item Set a~new state for a~statement \\(\verb|POST /statements/1?state=approved&user=Alice|).
\end{itemize}

All GET requests support an additional \verb|&state={state}| query parameter to filter by state.
This allows Wikidata developers to examine statements
that have already been updated (\emph{e.g.}, ``return 10 random rejected statements'')
for further analysis.
The default of this parameter is \verb|unapproved|.

In addition to these main request types, the REST API also provides endpoints
for importing and deleting datasets, as well as for retrieving the system's status.%
\footnote{The system's status is also presented in more convenient form at
\url{http://tools.wmflabs.org/wikidata-primary-sources/status.html}.}

All update activity around statements is logged in an activity log
for statistical purposes using the Wikidata user name of the user who performed the update.
Activity information can be displayed in an aggregated leader board form
(top users, top rejecters, trends, \emph{etc.}) to add aspects of gamification to the migration process.

\subsection{Front-end}

For the front-end, our objective was to model the migration process
as close to the natural Wikidata editing flow as possible.
Ideally, the user should barely notice the different underlying datasources
when navigating Wikidata.

To achieve this, we have in a~first step created a~Wikidata \emph{user script}.
Wikidata user scripts are part of the Wikidata tool chain%
\footnote{Wikidata tool chain:
\url{https://www.wikidata.org/wiki/Wikidata:Tools}}
and are created by users, but unlike gadgets---to be explained in the following---%
do not appear in a~user's preferences.
Instead, one has to add a~line of code into ones \texttt{common.js} file
in order to activate and use the script.
User scripts can be created by anyone and might not always be stable.

Once a~user script has matured, it can be converted into a~\emph{gadget}.
Gadgets are scripts that are likewise created by users,
but which can be simply enabled in user preferences under the section ``Gadgets''.
They can only be edited by administrators and are assumed to be stable.

With our front-end, we went through this maturation steps
and started with a~user script called \texttt{freebase2wikidata.js}%
\footnote{Initial user script:
\url{https://www.wikidata.org/wiki/User:Tomayac/freebase2wikidata.js}}
that we later converted into a~gadget.%
\footnote{Final gadget: \url{https://www.wikidata.org/wiki/MediaWiki:Gadget-PrimarySources.js}}
\autoref{fig:primary-sources-tool} illustrates the user interface in action.
Existing Wikidata statements are shown as usual,
Freebase facts that could potentially be imported into Wikidata are shown in light blue.
Instead of an ``edit'' button they feature two new buttons,
``approve reference'' and ``reject reference''
(\autoref{fig:barack-obama}) or ``approve claim'' and ``reject claim''
(\autoref{fig:cause-of-death}) respectively.

\subsection{Data Rollout}

In order to prepare the data for integration into the Primary Sources Tool,
we have created a~set of scripts which map the content
of the last Freebase dump to Wikidata statements.%
\footnote{The source code of theses scripts is also available under the terms of the Apache~2.0 license:
\url{https://github.com/google/freebase-wikidata-converter}}
The Primary Sources Tool was developed in parallel with the mapping scripts.
We first created a~small dataset of around 1 million statements
based on few select properties (\emph{e.g.}, \texttt{/people/person/birth\_place})
and deployed both the first dataset and a~basic version of the tool
in order to gather initial user feedback.

In the following months, we progressively added more statements to the tool's database back-end.
Thereby, we were able to slowly upscale the tool's database back-end
without risking back-end performance bottlenecks.
The process of adding data in small batches of 1~million to 2~million statements per batch
further allowed us to detect some potential issues in the front-end
and fix them before they became actual problems.
Another positive side-effect of this approach was that it allowed us to gather quick feedback
from the community about the already mapped data and thus to adapt the mapping scripts early
and correct minor issues with the data in the back-end.

Upon approval or rejection of either a~claim or a~reference,
the Wikidata item page in question reloads in order to reflect
the potentially changed state of the migration, as,
in case of incoming claims with references, freshly accepted claims 
can require incoming references to be attached elsewhere in the item.

\section{Statistics on the Migration}\label{sec:statistics-of-the-migration}

\begin{figure*}[h]
    \centering
    \begin{subfigure}[b]{1.0\columnwidth}
        \includegraphics[width=\textwidth]{img/freebase-projection-bw.pdf}
        \caption{Spatio-temporal distribution of Freebase.}
        \label{fig:freebase}
    \end{subfigure}
    \begin{subfigure}[b]{1.0\columnwidth}
        \includegraphics[width=\textwidth]{img/wikidata-projection-bw.pdf}
        \caption{Spatio-temporal distribution of Wikidata.}
        \label{fig:wikidata}
    \end{subfigure}
    \caption{Spatio-temporal distribution of Freebase and Wikidata items,
      the $x$ axis shows longitude degrees from 180°~E to 180°~W, the $y$ axis shows time
      from the year 0 to 2015, following a~power law ($y' = y^8$).}
    \label{fig:time-space}
\end{figure*}

\subsection{Quantitative Comparison}

In order to put the statistics in this section in context,
we first provide an overview of the size of the last dump of Freebase
from March 2015:

\begin{itemize}
  \setlength\itemsep{0em}
  \item 48~million topics
  \item 2,997~million triples
  \item 442~million ``useful'' facts%
  \item 68~million labels
\end{itemize}

``Useful'' facts are the triples that remain 
after filtering out all triples about IDs (\texttt{/type/object/id}),
types \linebreak (\texttt{/type/object/type}), labels, descriptions \linebreak
(\texttt{/type/object/name},
\texttt{/common/topic/description}, \emph{etc.}), \linebreak and triples removed
for licensing reasons (the latter were about 42~million, see \autoref{sec:licensing}).
The triples removed in this step are not considered to be potentially useful for
Wikidata, since IDs, labels, and descriptions are not handled through statements.

As a~direct comparison, here are the raw statistics for Wikidata as of August 2015.

\begin{itemize}
    \setlength\itemsep{0em}
    \item 14.5~million items
    \item 66~million statements
    \item 82~million labels
\end{itemize}

Why are these numbers so different?
Regarding the number of topics, Freebase has a~lot of topics about subjects
that do not match Wikidata's notability criteria.%
\footnote{Wikidata notability criteria:
\url{https://www.wikidata.org/wiki/Wikidata:Notability}}
For example, Freebase holds data about 11.3~million musical recordings,
8.3~million music canonical versions, 2.6~million ISBNs, \emph{etc.} that are not contained in Wikidata.

We also cannot compare the number of facts and the number of statements directly,
even considering the lower number of topics.
Freebase encodes its knowledge far more redundantly than Wikidata.
To illustrate this, properties in Freebase often have a~\emph{reverse} property
that is used in order to be able to traverse the Freebase graph easily in both directions.
For example, the property \texttt{/people/person/place\_of\_birth} has the corresponding
reverse property \linebreak \texttt{/location/location/people\_born\_here}
that encodes exactly the same semantic information.
Such reverse properties sometimes exist in Wikidata
(like \textit{``children''} for \textit{``father''} and \textit{``mother''}),
but are far less common and also not automatically handled.

Another point is that CVTs use a~lot of facts to convey the exact same information
that can be represented with a~single Wikidata statement.
For example, to encode that Barack Obama is the president
of the United States since January~20, 2009,
Freebase requires more than six facts as shown in \autoref{fig:cvt-obama}.
If we attempted to encode Wikidata statements as if they were Freebase facts, \emph{i.e.},
by removing sources, representing statements with qualifiers using CVTs,
and adding reverse properties, this would lead to a number of 110~million facts,
\emph{i.e.} an increase of 167\% over the raw number of statements.
The representation of Wikidata in RDF and its impact on the number of triples
is discussed in more detail in~\cite{reification-wikidata}.

In a~considerable amount of cases, Freebase also contains duplicate data.
For example, many cities in the United States have duplicate CVTs for the population,
where the only difference is the encoding of the source.

\subsection{Spatio-Temporal Comparison}

In order to get an idea of the coverage of the two knowledge bases,
we have further created a~spatio-temporal visualization of Freebase and Wikidata 
that is shown in \autoref{fig:time-space}.

We extracted the earliest date and the longitude
linked to each of the entities of the two knowledge bases.
We then propagated them to the connected entities that do not have such data.
To implement this propagation, we set as value (date or longitude)
the average of the values of the entities whose distance is $<10$
with a weight of $2^{-d}$ (with $d$~being the distance between the two entities).

We see that the two knowledge bases display the same basic patterns,
with their coverage mostly concentrated on the longitudes of Europe and North America.
There are no strong differences between them,
yet coverage of \mbox{\emph{post}-year-2000} entities seems stronger in Freebase.

Our theory is that this is
\emph{(i)}~due to the absence of a~lot of musical data in Wikidata that is present in Freebase,
and \emph{(ii)} due to the US population CVTs in Freebase.

The big horizontal lines in \autoref{fig:freebase} and even more in \autoref{fig:wikidata}
can most probably be traced back to places that do not have any linked date,
and in consequence inherit the foundation date of their country.
For Wikidata (\autoref{fig:wikidata}), two lines that are definitely caused by this effect
are the big line over the USA around 1780
and the one covering Russia around 850.

One possible conclusion from this visualization could be that Wikidata,
despite its smaller absolute size, appears to have
a~nearly as complete coverage of the world as Freebase.
Nonetheless, it is very challenging to find objective ways to compare the two databases
and good metrics of success for the migration.
We certainly can note that for data sets to be added to a knowledge bases,
the raw number of triples is not an adequate measure of quality.

\subsection{Raw Statistics}

From the Freebase dump, we have been able to create more than 17~million Wikidata claims
(\emph{i.e.}, statements without references), including 1.5~million IDs.
For that, we have used 31~million Freebase facts.
Including the references, we obtain 19.6~million statements, and, after removing duplicates
and facts already contained in Wikidata, we obtain 14~million new statements.

If all these statements were added to Wikidata,
we would see a~21\% increase of the number of statements in Wikidata.

This raises the question why these numbers are so low compared to the size of Freebase, given at
3~billion facts.
The main reason is that we have only mapped 4.56~million items to Wikidata,
\emph{i.e.}, only 9.5\% of Freebase's topics in total.
The mapped topics are the subject of only 64~million facts.
So even under otherwise ideal circumstances, we cannot map more than 64~million statements.
This assumes that we could map all reverse properties, which, in fact, we could not.
So leaving aside the not mapped topics, we have created a~statement for more than 24\% of the facts.

If we restrict ourselves to reviewed facts---a~set of
1.6~million human curated facts---we have far better results.
Of the human curated facts, 0.58~million facts have their subject mapped to Wikidata.
Based on our property mappings, 0.52~million of these facts
(\emph{i.e.}, 92\%) are converted to Wikidata statements.
Finally, 58\% of the statements created from reviewed fact are already in Wikidata,
allowing us to add 0.25~million new reviewed statements to Wikidata.

\section{Related Work}\label{sec:related-work}

For the challenges listed in \autoref{sec:challenges-of-the-migration}
there is a~body of work dealing with it. In many cases, the problems we have encountered in
this migration have turned out to be much simpler than the problems solved in existing research.

\textbf{Data licensing.}
Although many books and tutorials on publishing data on the Web of Data
(e.g.~\cite{linkeddatabook,linkeddatabestpractises}) mention the importance of licenses,
they are often added as an afterthought.
They have been preciously few dedicated publications on the
topic~\cite{licenses,licensius}, none of them written by lawyers.
Due to the lack of understanding surronding these topics, we have been
cautious and chose to simply remove any facts from the migrating data
where the situation is in doubt.

\textbf{Schema mappings.}
There has been extensive research on the topic of mapping the structures
of two databases or knowledge bases
(some surveys are~\cite{mappingsurvey1,mappingsurvey2,mappingsurvey3}).
The current work is, in many technical respects, rather simple:
we are looking for a~one-directional mapping (since the case of data
flowing from Wikidata back to Freebase is not relevant),
a~one-off data flow (since Freebase is now read-only),
mostly individually evaluated by human contributors
(therefore not requiring a perfect mapping).
The effort most similar to ours is the DBpedia Mapping Wiki~\cite{lehmann2014dbpedia}.
The DBpedia community is manually writing and maintaining the mapping from
the DBpedia extraction process (working on Wikipedia) to the DBpedia ontology.
The main difference is that the DBpedia mapping is running continuously
and does not allow for point-fixes, thus necessitating a higher quality of the mapping,
whereas for the Freebase to Wikidata mapping every single results is expected to be curated.

\textbf{Entity mappings.}
Entity mapping (often called entity linking) deals with finding the objects
in several sources that refer to the same entity in the domain of discourse.
There is extensive work in this area (e.g.~\cite{entitylinkingsurvey1,entitylinkingsurvey2})
but we avoided most of the challenges by relying on the simple fact that
both Freebase and Wikidata have historically been bootstrapped by using Wikipedia.
Wikipedia is often used as a benchmark for entity mapping systems~\cite{entitylinkingwikipedia}.
As described in \autoref{subsec:topic-mappings} this lead to sufficiently good results for the head,
and we think it would be surprising if further effort in this area would lead to reasonable gains.

\textbf{Data maintenance.}
The Wikidata requirements~\cite{wikidatareqs} states that the data uploaded to Wikidata
must not overwhelm the community.
Research in understanding how systems for collaborative knowledge creation are impacted
by events like this data migration are still in its early stages~\cite{moskaliuk2012collaborative},
in particular for structured knowledge~\cite{horridge2014webprotege}.
Most of the research is focused on Wikipedia~\cite{flock2015towards},
which is understandable considering the availability of its data sets,
in particular the whole edit history~\cite{schindler2011introducing}
and the availability of tools for working with Wikipedia~\cite{milne2013open}.

%Should also cover references (kvault) and quality.

\section{Future Work and Conclusion}\label{sec:future-work-and-conclusion}

The largest gains for the migration can be achieved
by extending the mapping to more Freebase topics.
A~possible way of realizing this is to create a~user interface%
---possibly leveraging elements of gamification---%
that would allow users to create new Wikidata items for a~suggested topic
or to add a~mapping for them to an existing Wikidata item.
In order to suggest interesting topics to add to Wikidata,
we could rank the not mapped yet topics per number of incoming links
from already mapped topics and filter less interesting types like ISBNs.
Another area for improvement is to upload high quality datasets using a~bot,
like the reviewed facts or some sets for external IDs,
in order to speed up the integration of Freebase content into Wikidata.
We have already started to upload simple reviewed facts about humans
like birth date, death place or gender using a~bot.
We have also begun to import labels that are in Freebase and are absent from Wikidata.
In the 17.2M~Freebase labels for mapped topics, only 0.9M, \emph{i.e.}, 5\%, are missing in Wikidata.

Concluding, in a~fairly short amount of time, we have been able
to provide to the Wikidata community more than\linebreak 14M~new Wikidata statements
using a~novel customizable and generalizable software solution called Primary Sources Tool
that is well integrated into the Wikidata user interface.
The effort needed to map two fairly different knowledge bases has also been a~good occasion
to highlight the difficulty of having good metrics to measure the size
and in consequence the ``value'' of such collaborative knowledge bases.
We envision that this project now with Freebase, and for even more datasets in the future,
will facilitate to increase the completeness of Wikidata.

\bibliographystyle{abbrv}
\bibliography{biblio}

\balancecolumns
\end{document}